---
title: "Bayesian Learning - Lab 03"
author: "Lakshidaa Saigiridharan (laksa656) and Maximilian Pfundstein (maxpf364)"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: false
    number_sections: false
  pdf_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
```

# Normal Model, Mixture of Normal Model with Semi-Conjugate Prior

**Exercise:** The data `rainfall.dat` consist of daily records, from the beginning of 1948 to the end of 1983, of precipitation (rain or snow in units of 1 inch, and records of zero 100 precipitation are excluded) at Snoqualmie Falls, Washington. Analyze the data using the following two models.

a) Assume the daily precipitation ${y_1,...,y_n}$ are independent normally distributed, $y_1,...,y_n|\mu,\sigma^2 \sim \mathcal{N}(\mu,\sigma^2)$ where both $\mu$ and $\sigma^2$ are unknown. Let $\mu \sim \mathcal{N}(\mu_0,\tau_0^2)$ independently of $\sigma^2 \sim Inv-\chi^2(\nu_0, \sigma_0^2)$

    - Implement (code!) a Gibbs sampler that simulates from the joint posterior $p(\mu,\sigma^2|y_1,...,y_n)$. The full conditional posteriors are given on the slides from Lecture 7.
    - Analyze the daily precipitation using your Gibbs sampler in (a)-i. Evaluate the convergence of the Gibbs sampler by suitable graphical methods, for example by plotting the trajectories of the sampled Markov chains.

b) Let us now instead assume that the daily precipitation ${y_1,...,y_n}$ follow an iid two-component **mixture of normals** model:

$$p(y_i|\mu,\sigma^2,\pi) = \pi \mathcal{N}(y_i|\mu_1, \sigma_1^2) + (1 - \pi) \mathcal{N}(y_i|\mu_2, \sigma_2^2)$$
where

$$\mu = (\mu_1, \mu_2) \; \text{and} \; \sigma^2 = (\sigma_1^2, \sigma_2^2)$$

Use the Gibbs sampling data augmentation algorithm in `NormalMixtureGibbs.R` (available under Lecture 7 on the course page) to analyze the daily precipitation data. Set the prior hyperparameters suitably. Evaluate the convergence of the sampler.

c) Plot the following densities in one figure:
    - A histogram or kernel density estimate of the data.
    - Normal density $\mathcal{N}(\mu, \sigma^2)$ in (a)
    - Mixture of normals density $p(y_i|\mu,\sigma^2,\pi)$ in (b)
    
Use the posterior mean value for all the parameters.

## Normal Model

### Gibbs Sampler Implementation

First we load the data and set the parameters for our prior.

```{r}

################################################################################
# Exercise 1.a i)
################################################################################

rainfall_data = read.table("data/rainfall.dat", header=FALSE)

# Prior Parameters (slides page 15)
# These are basically random, guessed or from the data
# mu
mu_0 = 0
tau_sq_0 = 1

# sigma_sq
nu_0 = 1
sigma_sq_0 = 1 # sigma is 1/nu_0

```

This is a custom function to sample from the inverse chi squared distribution.

```{r}

crinvchisq = function(n, df, s_sq) {
  samples = rchisq(n, df)
  # These are draws from the inverse chi squared
  sigma_sq = (df - 1) * s_sq / samples
  return(sigma_sq)
}

```

This is the Gibbs Sampler, which takes the number of draws, a default $\sigma$ (as both $\sigma$ and $\mu$ depend on each other we need to start somewhere) and some more parameters to calculate the posterior parameters.

```{r}

gibbs_sample = function(nDraws, data, default_sigma, tau_sq_0, mu_0, nu_0, sigma_sq_0) {
  
  # Posterior Parameters (Taken from lecture 2 slide 4)
  n = length(data)
  mu_n = mean(data) + mu_0
  nu_n = nu_0 + n
  default_sigma_sq = default_sigma^2
  
  # To store all iterative results
  values_df = data.frame(matrix(NA, nrow = nDraws, ncol = 2))
  
  # To save current iterative results
  values = list(mu = NaN, sigma_sq = default_sigma_sq)
  
  # As mu depends on sigma and sigma depends on mu, we need one initial value to start.
  # In this implementation the default value to start with is sigma_sq.
  
  for (i in 1:nDraws) {
    tau_sq_n = 1 / ((n/values$sigma) + (1/tau_sq_0))
    values$mu = rnorm(1, mu_n, sqrt(tau_sq_n))
    values$sigma_sq = crinvchisq(1, nu_n, (nu_0*sigma_sq_0 + sum((data - values$mu)^2))/(n + nu_0))
    values_df[i,] = values
  }
  
  return(values_df)
}

```

Now we call our Gibbs Sampler and print the draws.

```{r}

res = gibbs_sample(nDraws = 1000,
                   data = rainfall_data$V1, 
                   default_sigma = 40,
                   tau_sq_0 = tau_sq_0,
                   mu_0 = mu_0,
                   nu_0 = nu_0,
                   sigma_sq_0 = sigma_sq_0)

kable(head(res))

```

### Analysing the Precipitation

This is the traceplot for $\mu$ including the trailing mean. We can see that the mean converges.

```{r, echo = FALSE}

################################################################################
# Exercise 1.a ii)
################################################################################

plotdf = data.frame(1:nrow(res), res, cumsum(res$X1)/(1:nrow(res)),
                    cumsum(res$X2)/(1:nrow(res)))
colnames(plotdf) = c("index", "mu", "sigma_sq", "mu_trailing_mean",
                     "sigma_sq_trailing_mean")

ggplot(plotdf) +
  geom_line(aes(x = index, y = mu), color = "#C70039") +
  geom_line(aes(x = index, y = mu_trailing_mean), color = "#000000") +
  labs(title = "Traceplot For Mu", y = "mu",
  x = "Iteration", color = "Legend") +
  theme_minimal()

```

The same for $\sigma^2$.

```{r, echo = FALSE}

ggplot(plotdf) +
  geom_line(aes(x = index, y = sigma_sq), color = "#C70039") +
  geom_line(aes(x = index, y = sigma_sq_trailing_mean), color = "#000000") +
  labs(title = "Traceplot For Mu", y = "mu",
  x = "Iteration", color = "Legend") +
  theme_minimal()

```

## Mixture Normal Model

```{r}

################################################################################
# Exercise 1.b)
################################################################################

# This is the script provied in 'NormalMixtureGibbs.R' with minor ajustments.

# Estimating a simple mixture of normals
# Author: Mattias Villani, IDA, Linkoping University. http://mattiasvillani.com

##########    BEGIN USER INPUT #################
# Data options
x <- as.matrix(rainfall_data$V1)

# Model options
## COMMENT: As the given model has two guassians we set this to two.
nComp <- 2    # Number of mixture components

# Prior options
alpha <- 10*rep(1,nComp) # Dirichlet(alpha)
muPrior <- rep(0,nComp) # Prior mean of mu
tau2Prior <- rep(10,nComp) # Prior std of mu
sigma2_0 <- rep(var(x),nComp) # s20 (best guess of sigma2)
nu0 <- rep(4,nComp) # degrees of freedom for prior on sigma2

# MCMC options
nIter <- 100 # Number of Gibbs sampling draws

# Plotting options
plotFit <- TRUE
lineColors <- c("blue", "green", "magenta", 'yellow')
# We removed the sleep time
sleepTime <- 0.1 # Adding sleep time between iterations for plotting
################   END USER INPUT ###############

###### Defining a function that simulates from the 
rScaledInvChi2 <- function(n, df, scale){
  return((df*scale)/rchisq(n,df=df))
}

####### Defining a function that simulates from a Dirichlet distribution
rDirichlet <- function(param){
  nCat <- length(param)
  piDraws <- matrix(NA,nCat,1)
  for (j in 1:nCat){
    piDraws[j] <- rgamma(1,param[j],1)
  }
  piDraws = piDraws/sum(piDraws) # Diving every column of piDraws by the sum of the elements in that column.
  return(piDraws)
}

# Simple function that converts between two different representations of the mixture allocation
S2alloc <- function(S){
  n <- dim(S)[1]
  alloc <- rep(0,n)
  for (i in 1:n){
    alloc[i] <- which(S[i,] == 1)
  }
  return(alloc)
}

# Initial value for the MCMC
nObs <- length(x)
S <- t(rmultinom(nObs, size = 1 , prob = rep(1/nComp,nComp))) # nObs-by-nComp matrix with component allocations.
mu <- quantile(x, probs = seq(0,1,length = nComp))
sigma2 <- rep(var(x),nComp)
probObsInComp <- rep(NA, nComp)

# Setting up the plot
xGrid <- seq(min(x)-1*apply(x,2,sd),max(x)+1*apply(x,2,sd),length = 100)
xGridMin <- min(xGrid)
xGridMax <- max(xGrid)
mixDensMean <- rep(0,length(xGrid))
effIterCount <- 0
ylim <- c(0,2*max(hist(x)$density))

# Added for stroring the draws and print the plots
all_mu = matrix(rep(NaN, nIter*nComp), nrow = nIter)
all_sigma_sq = matrix(rep(NaN, nIter*nComp), nrow = nIter)

for (k in 1:nIter){
  message(paste('Iteration number:',k))
  alloc <- S2alloc(S) # Just a function that converts between different representations of the group allocations
  nAlloc <- colSums(S)
  print(nAlloc)
  # Update components probabilities
  pi <- rDirichlet(alpha + nAlloc)
  
  # Update mu's
  for (j in 1:nComp){
    precPrior <- 1/tau2Prior[j]
    precData <- nAlloc[j]/sigma2[j]
    precPost <- precPrior + precData
    wPrior <- precPrior/precPost
    muPost <- wPrior*muPrior + (1-wPrior)*mean(x[alloc == j])
    tau2Post <- 1/precPost
    mu[j] <- rnorm(1, mean = muPost, sd = sqrt(tau2Post))
  }
  
  # Update sigma2's
  for (j in 1:nComp){
    sigma2[j] <- rScaledInvChi2(1, df = nu0[j] + nAlloc[j], scale = (nu0[j]*sigma2_0[j] + sum((x[alloc == j] - mu[j])^2))/(nu0[j] + nAlloc[j]))
  }
  
    all_mu[k,] = mu
    all_sigma_sq[k,] = sigma2
  
  # Update allocation
  for (i in 1:nObs){
    for (j in 1:nComp){
      probObsInComp[j] <- pi[j]*dnorm(x[i], mean = mu[j], sd = sqrt(sigma2[j]))
    }
    S[i,] <- t(rmultinom(1, size = 1 , prob = probObsInComp/sum(probObsInComp)))
  }
  
  # Printing the fitted density against data histogram
  if (plotFit && (k == nIter)){
    effIterCount <- effIterCount + 1
    hist(x, breaks = 20, freq = FALSE, xlim = c(xGridMin,xGridMax), main = paste("Iteration number",k), ylim = ylim)
    mixDens <- rep(0,length(xGrid))
    components <- c()
    for (j in 1:nComp){
      compDens <- dnorm(xGrid,mu[j],sd = sqrt(sigma2[j]))
      mixDens <- mixDens + pi[j]*compDens
      lines(xGrid, compDens, type = "l", lwd = 2, col = lineColors[j])
      components[j] <- paste("Component ",j)
    }
    mixDensMean <- ((effIterCount-1)*mixDensMean + mixDens)/effIterCount
    
    lines(xGrid, mixDens, type = "l", lty = 2, lwd = 3, col = 'red')
    legend("topleft", box.lty = 1, legend = c("Data histogram",components, 'Mixture'), 
           col = c("black",lineColors[1:nComp], 'red'), lwd = 2)
    #Sys.sleep(sleepTime)
  }
}

hist(x, breaks = 20, freq = FALSE, xlim = c(xGridMin,xGridMax), main = "Final fitted density")
lines(xGrid, mixDensMean, type = "l", lwd = 2, lty = 4, col = "red")
lines(xGrid, dnorm(xGrid, mean = mean(x), sd = apply(x,2,sd)), type = "l", lwd = 2, col = "blue")
legend("topright", box.lty = 1, legend = c("Data histogram","Mixture density","Normal density"), col=c("black","red","blue"), lwd = 2)

#########################    Helper functions    ##############################################


```

This is the traceplot for $\mu$.

```{r, echo = FALSE}

mu_plot_dataframe = data.frame(1:nIter, all_mu, cumsum(all_mu[,1])/(1:nrow(all_mu)),
                               cumsum(all_mu[,2])/(1:nrow(all_mu)))
colnames(mu_plot_dataframe) = c("index", "mu_1", "mu_2",
                                "mu_1_trailing_mean", "mu_2_trailing_mean")

ggplot(mu_plot_dataframe) +
  geom_line(aes(x = index, y = mu_1), color = "#C70039") +
  geom_line(aes(x = index, y = mu_1_trailing_mean), color = "#000000") +
  geom_line(aes(x = index, y = mu_2), color = "#6091EC") +
  geom_line(aes(x = index, y = mu_2_trailing_mean), color = "#000000") +
  labs(title = "Traceplot for mu_1 and mu_2", y = "mu",
  x = "Iteration", color = "Legend") +
  theme_minimal()

```

This is the traceplot for $\sigma$.

```{r, echo = FALSE}

sigma_sq_plot_dataframe = data.frame(1:nIter, all_sigma_sq, cumsum(all_sigma_sq[,1])/(1:nrow(all_sigma_sq)),
                               cumsum(all_sigma_sq[,2])/(1:nrow(all_sigma_sq)))
colnames(sigma_sq_plot_dataframe) = c("index", "sigma_sq_1", "sigma_sq_2",
                                "sigma_sq_1_trailing_mean", "sigma_sq_2_trailing_mean")

ggplot(sigma_sq_plot_dataframe) +
  geom_line(aes(x = index, y = sigma_sq_1), color = "#C70039") +
  geom_line(aes(x = index, y = sigma_sq_1_trailing_mean), color = "#000000") +
  geom_line(aes(x = index, y = sigma_sq_2), color = "#6091EC") +
  geom_line(aes(x = index, y = sigma_sq_2_trailing_mean), color = "#000000") +
  labs(title = "Traceplot for sigma_sq_1 and sigma_sq_2", y = "sigma_sq",
  x = "Iteration", color = "Legend") +
  theme_minimal()

```

**Parameters of the Prior:** Without looking at the data we have no clear idea of the priors should look like. We can assume that probably the mean is greater than 0, but about the variance or $\alpha$ we do not really have an idea. That's why we leave the parameters as they are. The default implementation actially takes the variance of the data for the prior, where we think that's not 100 percent appropriate but it depends on the problems one tackles in the real world of course. In addition it shouldn't matter that much because after the burn-in period the prior shouldn't matter any more to discover the posterior density function.

**Convergence:** As we can see from the different trace plots the parameters actually converge the same as in the previous exercise. More iterations will make the result more accurate but we set it to a low amount of iterations to have a better zoom on the start of the process to observe the burn-in period.

## Graphical Comparison

# Metropolis Random Walk for Poisson Regression

**Exercise:** Consider the following Poisson regression model

$$y_i|\beta \sim \text{Poisson}\left[ \text{exp}(x_i^T\beta)  \right], i = 1,...,n$$

where $y_i$ is the count for the ith observation in the sample and $x_i$ is the p-dimensional vector with covariate observations for the ith observation. Use the data set `eBayNumberOfBidderData.dat`. This dataset contains observations from 1000 eBay auctions of coins. The response variable is **nBids** and records the number of bids in each auction. The remaining variables are features/covariates (**x**):

- **Const** (for the intercept)
- **PowerSeller** (is the seller selling large volumes on eBay?)
- **VerifyID** (is the seller verified by eBay?)
- **Sealed** (was the coin sold sealed in never opened envelope?)
- **MinBlem** (did the coin have a minor defect?)
- **MajBlem** (a major defect?)
- **LargNeg** (did the seller get a lot of negative feedback from customers?)
- **LogBook** (logarithm of the coins book value according to expert sellers. Stan- dardized)
- **MinBidShare** (a variable that measures ratio of the minimum selling price (starting price) to the book value. Standardized).

a) Obtain the maximum likelihood estimator of $\beta$ in the Poisson regression model for the eBay data [Hint: `glm.R`, don’t forget that `glm()` adds its own intercept so don’t input the covariate Const]. Which covariates are significant?

b) Let’s now do a Bayesian analysis of the Poisson regression. Let the prior be $\beta \sim \mathcal{N}[\mathbf{0}, 100 \cdot (X^TX)^{-1}]$ where **X** is the $n \times p$ covariate matrix. This is a commonly used prior which is called Zellner’s g-prior. Assume first that the posterior density is approximately multivariate normal:

$$\beta|y \sim \mathcal{N}[\tilde{\beta}, J_y^{-1}(\tilde{\beta})],$$

where $\tilde{\beta}$ is the posterior mode and $J_y(\tilde{\beta})$ is the negative Hessian at the posterior mode. $\tilde{\beta}$ and $J_y(\tilde{\beta})$ an be obtained by numerical optimization (`optim.R`) exactly like you already did for the logistic regression in Lab 2 (but with the log posterior function replaced by the corresponding one for the Poisson model, which you have to code up.).

c) Now, let’s simulate from the actual posterior of $\beta$ using the Metropolis algorithm and compare with the approximate results in b). Program a general function that uses the Metropolis algorithm to generate random draws from an arbitrary posterior density. In order to show that it is a general function for any model, I will denote the vector of model parameters by $\theta$. Let the proposal density be the multivariate normal density mentioned in Lecture 8 (random walk Metropolis):

$$\theta_p|\theta^{(i-1)} \sim \mathcal{N}(\theta^{(i-1)}, c \cdot \Sigma),$$

where $\Sigma = J_y^{-1}(\tilde{\beta})$ obtained in b). The value c is a tuning parameter and should be an input to your Metropolis function. The user of your Metropolis function should be able to supply her own posterior density function, not necessarily for the Poisson regression, and still be able to use your Metropolis function. This is not so straightforward, unless you have come across *function objects* in R and the triple dot ( ... ) wildcard argument. I have posted a note (`HowToCodeRWM.pdf`) on the course web page that describes how to do thisin R.

Now, use your new Metropolis function to sample from the posterior of $\beta$ in the Poisson regression for the eBay dataset. Assess MCMC convergence by graphical methods.

## Maximum Likelihood Estimator

## Bayesian Analysis of the Poisson Regression

## Simulate from the Actual Posterior Using the Metropolis Algorithm

# Source Code

```{r, ref.label=knitr::all_labels(), echo = TRUE, eval = FALSE, results = 'show'}

```